\topmatter{Acknowledgments}

First, I would like to thank my family: my parents Xueyi Chen and Cuilan Zhou, for giving birth to me and passing their value on hard work and education to me. My sister Li Chen and my brother Yuwei Chen, who have continuously encouraged me to pursuit my dreams. 

I would like to express my sincere gratitude to my advisor Ewa Deelman for the continuous support of my Ph.D. study and research. She provided me with excellent guidance, encouraged me investigate my own ideas, connected me with many experts in different disciplines and corrected my papers, posters, articles, presentations and thesis. I could not have imaged having a better advisor and mentor for my Ph.D. study. 

Besides my advisor, I am deeply grateful to my committee members, Bob Lucas, Ann Chervenak, Aiichiro Nakano and Viktor Prasanna, who have all given me invaluable guidance and advice. I took Bob Lucas' Parallel Programming course and he helped me build a whole picture of the parallel and distributed world. I remember my first project at ISI was with Ann Chervenak and Muhammad Ali Amer. They helped me build the very skills for my study and research and I am grateful to them for enlightening me the first glance of research.

I am fortunate to have many good friends and colleagues at ISI. I would like to thank Karan Vahi, Gaurang Mehta, Mats Rynge. They helped me learn many softwares and tools I needed in my research. Their patience, enthusiasm and immense knowledge helped me overcome many problems I have met in practice. 

I would like to offer my special thanks to Gideon Juve, who have been a wise counselor and a great source of ideas and discussion.  Gideon have continuously offered me his great advice not only on my research and study, but also on my career planning and selection. 

I am fortunate to meet Rafael Ferreira da Silva at the last year of my Phd, who provided me a sounded support of my research. 
%With similar research interest, we often exchanged our ideas about our works. 
And I would like to offer my special appreciation to Rafael for all of his work on the research described in Chapter 5 and Chapter 6. Rafael helped me finalize the algorithms and metrics used for balanced task clustering and fault tolerant clustering. 

I thank my other labmates in Pegasus group: Rajiv Mayani, Prasanth Thomas, Fabio Silva, Jens V\"{o}ckler and Idafen Santana P\'{e}rez for the stimulating discussions and for all the fun we have had in the last five years. 

I have greatly benefited from three great European scientists: Rizos Sakellariou, Thomas Fahringer and Radu Prodan. The discussion with them have been very insightful and inspiring. Radu Prodan's work on overhead analysis inspired my work on task clustering. Rizos Sakellariou helped me finalize the algorithms and metrics used for balanced task clustering. Thomas Fahringer worked with me on the chapter of fault tolerant clustering and provided his insightful ideas. 

I thank my friends in the networking and AI division: Congxing Cai, Lin Quan, Zi Hu, Xun Fan, Xue Cai, Chengjie Zhang, Xiyue Deng, Hao Shi and Lihang Zhao for all the fun we have had at ISI and their advice on my thesis. 

Finally, I would like to thank everyone at ISI, the Viterbi School of Engineering, and the Department of Computer Science, who have helped me reach this point: Larry Godinez, Liszl DeLeon, Ruth Heron, Raphael Bolze, Rubing Duan, Claire Bono and Mei-Hui Su. 

